---
title: "BA_Homework_6"
author: "Ali Kulunyar"
date: "November 26, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Relation between Oil, Gold and SP 500
cff plots show that Gold, Oil and SP500 are correlated within each other. The point where h=-0.9 gold and oil has the strongest positive correlation. Moreover, the point where h=-0.09, oil and sp500 has the strongest anti correlation. It is also same for gold and sp500. Additionaly, combined time-series graphs support this claim visually.

## Investment Opportunity 
Given no experience in finance related predictions, Oil and Gold have ups and downs for past 5 years. Futhermore, the forecasts are suggesting a decrease in oil and gold prices in the next year(2017). However, for SP500 prices, news are all good. ARIMA forecasted that the prices will be up around 2158 dollars. That is why we could expect more return by investing in SP500.


##Data Import, Date Modification
```{r ts}
getwd()
setwd('/Users/mkulunyar/Dropbox/NYU/Fall16/BusinessAnalytics/HW_6')

#Data Export
#investing.com
oil<-read.csv('new_oil_daily.csv', header=TRUE, stringsAsFactors=FALSE)
gold<-read.csv('new_gold_monthly.csv', header=TRUE, stringsAsFactors=FALSE)
sp500<-read.csv('new_sp500_daily.csv', header=TRUE, stringsAsFactors=FALSE)

#DATE
#Date Modification for Oil
head(oil)
oil$Date <- as.Date(oil$Date, "%d-%B-%y")
oil_sorted<-oil[order(oil$Date, decreasing = FALSE),]
head(oil_sorted)
#Date Modification for sp500
head(sp500)
sp500$Date <- as.Date(sp500$Date, "%d-%B-%y")
sp500_sorted<-sp500[order(sp500$Date, decreasing = FALSE),]
head(sp500_sorted)
#Date Modification for gold
head(gold)
gold$Date <- as.Date(gold$Date, "%y-%B-%d")
gold_sorted<-gold[order(gold$Date, decreasing = FALSE),]
head(gold_sorted)

#Time Series
#Oil
#Daily
summary(oil_sorted$Price)
Crude_Oil.ts <- ts(oil_sorted$Price, frequency = 365, start=c(2011))
plot(Crude_Oil.ts)
#Monthly
head(oil_sorted)
oil_sorted$MonthYear<-format(oil_sorted$Date, format="%Y-%m")
oil_monthly <- aggregate(oil_sorted$Price ~ MonthYear, oil_sorted, mean, na.rm=TRUE)
oil_monthly
oil_monthly.ts<-ts(oil_monthly$`oil_sorted$Price`, frequency = 12, start = c(2011))
plot(oil_monthly.ts)
#Gold
summary(gold_sorted$Price)
Gold.ts <- ts(gold_sorted$Price, frequency = 12, start=c(2011))
plot(Gold.ts)
#sp500
#Daily
summary(sp500_sorted$Price)
sp500.ts <- ts(sp500_sorted$Price, frequency = 365, start=c(2011))
plot(sp500.ts)
#Monthly
head(sp500_sorted)
sp500_sorted$MonthYear<-format(sp500_sorted$Date, format="%Y-%m")
sp500_monthly <- aggregate(sp500_sorted$Price ~ MonthYear, sp500_sorted, mean, na.rm=TRUE)
sp500_monthly
sp500_monthly.ts<-ts(sp500_monthly$`sp500_sorted$Price`, frequency = 12, start = c(2011))
plot(sp500_monthly.ts)
```


##Decomposition, Combining Time Series by using Oil and SP500 monthly
```{r ts2}
#Decomposition
#Gold
Gold.ts.d<-decompose(Gold.ts)
plot(Gold.ts.d)
#Oil
Crude_Oil.ts.d<-decompose(oil_monthly.ts)
#sp500
sp500.ts.d<-decompose(sp500_monthly.ts)
plot(sp500.ts.d)

#Combining Oil, Gold and sp500
library(xts)
#Oil-Gold
oil_gold <- cbind(oil_monthly.ts, Gold.ts)
plot(oil_gold)
#gold-sp500
gold_sp500 <- cbind(Gold.ts, sp500_monthly.ts)
plot(gold_sp500)
#oil-sp500
oil_sp500 <- cbind(oil_monthly.ts, sp500_monthly.ts)
plot(oil_sp500)
#Some Visuals
ts.plot(oil_monthly.ts, Gold.ts,gpars= list(col=rainbow(10)))
ts.plot(oil_monthly.ts, sp500_monthly.ts,gpars= list(col=rainbow(10)))
ts.plot(Gold.ts, sp500_monthly.ts,gpars= list(col=rainbow(10)))
```

##Cross-Correlation (cff graphs)
```{r ts4}
require(graphics)
#Cross Correlation Plots, Lets see what effects oil prices
ccf(Gold.ts, oil_monthly.ts, ylab = "cross-correlation")
ccf(sp500_monthly.ts, oil_monthly.ts, ylab = "cross-correlation")
ccf(sp500_monthly.ts, Gold.ts, ylab = "cross-correlation")
```

##Smoothing and Forecasting using Holt's and ARIMA for a horizon of 12 months
```{r ts3}
#oil
#smoothing
oil_monthly.holt.T <- HoltWinters(oil_monthly.ts, gamma=TRUE)
oil_monthly.holt.T
plot(oil_monthly.holt.T)
oil_monthly.holt.F <- HoltWinters(oil_monthly.ts, gamma=FALSE)
oil_monthly.holt.F
plot(oil_monthly.holt.F)
#forecasting, holt, horizon is 12 months
library(forecast)
oil_monthly.forecasts <- forecast.HoltWinters(oil_monthly.holt.T, h=12)  
plot.forecast(oil_monthly.forecasts)
oil_monthly.forecasts <- forecast.HoltWinters(oil_monthly.holt.F, h=12)  
plot.forecast(oil_monthly.forecasts)
#ARIMA
oil_monthly.arima<-arima(oil_monthly.ts, c(0,0,0))  
oil_monthly.arima.forecasts <- forecast.Arima(oil_monthly.arima, h=12)
oil_monthly.arima.forecasts
plot(oil_monthly.arima.forecasts)

oil_monthly.arima<-arima(oil_monthly.ts, c(0,1,0))  
oil_monthly.arima.forecasts <- forecast.Arima(oil_monthly.arima, h=12)
oil_monthly.arima.forecasts
plot(oil_monthly.arima.forecasts)

oil_monthly.arima<-arima(oil_monthly.ts, c(1,1,0))  
oil_monthly.arima.forecasts <- forecast.Arima(oil_monthly.arima, h=12)
oil_monthly.arima.forecasts
plot(oil_monthly.arima.forecasts)

#gold
#smoothing
Gold.holt.T <- HoltWinters(Gold.ts, gamma=TRUE)
Gold.holt.T
plot(Gold.holt.T)
Gold.holt.F <- HoltWinters(Gold.ts, gamma=FALSE)
Gold.holt.F
plot(Gold.holt.F)
#forecasting, holt, horizon is 12 months
library(forecast)
Gold.forecasts <- forecast.HoltWinters(Gold.holt.T, h=12)  
plot.forecast(Gold.forecasts)
Gold.forecasts <- forecast.HoltWinters(Gold.holt.F, h=12)  
plot.forecast(Gold.forecasts)
#ARIMA
Gold.arima<-arima(Gold.ts, c(0,0,0))  
Gold.arima.forecasts <- forecast.Arima(Gold.arima, h=12)
Gold.arima.forecasts
plot(Gold.arima.forecasts)

Gold.arima<-arima(Gold.ts, c(0,1,0))  
Gold.arima.forecasts <- forecast.Arima(Gold.arima, h=12)
Gold.arima.forecasts
plot(Gold.arima.forecasts)

Gold.arima<-arima(Gold.ts, c(1,1,0))  
Gold.arima.forecasts <- forecast.Arima(Gold.arima, h=12)
Gold.arima.forecasts
plot(Gold.arima.forecasts)


#sp500
#smoothing
sp500_monthly.holt.T <- HoltWinters(sp500_monthly.ts, gamma=TRUE)
sp500_monthly.holt.T
plot(sp500_monthly.holt.T)
sp500_monthly.holt.F <- HoltWinters(sp500_monthly.ts, gamma=FALSE)
sp500_monthly.holt.F
plot(sp500_monthly.holt.F)
#forecasting, holt, horizon is 12 months
library(forecast)
sp500_monthly.forecasts <- forecast.HoltWinters(sp500_monthly.holt.T, h=12)  
plot.forecast(sp500_monthly.forecasts)
sp500_monthly.forecasts <- forecast.HoltWinters(sp500_monthly.holt.F, h=12)  
plot.forecast(sp500_monthly.forecasts)
#ARIMA
sp500_monthly.arima<-arima(sp500_monthly.ts, c(0,0,0))  
sp500_monthly.arima.forecasts <- forecast.Arima(sp500_monthly.arima, h=12)
sp500_monthly.arima.forecasts
plot(sp500_monthly.arima.forecasts)

sp500_monthly.arima<-arima(sp500_monthly.ts, c(0,1,0))  
sp500_monthly.arima.forecasts <- forecast.Arima(sp500_monthly.arima, h=12)
sp500_monthly.arima.forecasts
plot(sp500_monthly.arima.forecasts)

sp500_monthly.arima<-arima(sp500_monthly.ts, c(1,1,0))  
sp500_monthly.arima.forecasts <- forecast.Arima(sp500_monthly.arima, h=12)
sp500_monthly.arima.forecasts
plot(sp500_monthly.arima.forecasts)
```
